Developing data load and processing

### 2020-08-17 dev/data_preprocessing.Rmd
Translation for preprocessing

```{r}
i <- 2
file_relative <- raw_data_map$File[i]
hospital_id <- raw_data_map$HospitalID[i]

@@

patient_col_types <- c("text","numeric","text","date","date", #ID - Date of admission
                                   rep("text", 12), #Transferred from - Smoking
                                   rep("numeric", 5), # BMI - Albumin
                                   "text", "text", "date", "text", "text") #Discontinued medication - Transferred to

patient_data_raw <- readxl::read_excel(file, sheet = "Patients", range = "A2:AA100", na = "NA", col_types = patient_col_types)

@@

logical_column <- function(x) {
  if(any( !is.na(x) && !(tolower(x) %in% c("yes","no"))) ) {
    stop("Invalid logical column contents")
  }
  tolower(x) == "yes"
}

@@

patient_data_typed <- patient_data_raw_usable %>% 
  transmute(
    hospital_id = hospital_id, 
    patient_id = `Patient ID`,
    age = Age, 
    sex = factor(Sex, levels = c("M","F")), 
    symptom_onset = lubridate::as_date(`Date of symptom onset`), 
    admission = lubridate::as_date(`Date of admission`), 
    transferred_from = `Transferred from`, 
    admitted_for_covid = logical_column(`Admitted for Covid`), 
    ischemic_heart_disease = logical_column(`Ischemic Heart Disease`),
    has_hypertension_drugs = tolower(`Hypertension drugs`) != "no",
    n_hypertension_drugs = if_else(has_hypertension_drugs,
                                   as.integer(if_else(has_hypertension_drugs, `Hypertension drugs`, NA_character_)), #Inner if_else to avoid warning
                                    0L),
    heart_failure = logical_column(`Heart failure`),
    COPD = logical_column(COPD),
    asthma = logical_column(`Asthma Bronchiale`),
    other_lung_disease = logical_column(`Other lung disease`),
    diabetes = logical_column(Diabetes),
    renal_disease = logical_column(`Renal Disease`),
    liver_disease = logical_column(`Liver Disease`),
    smoking = logical_column(Smoking),
    BMI = BMI,
    NYHA = NYHA,
    creatinin = `Creatinin [μmol/L]`,
    pt_inr = `PT INR (Quick)`,
    albumin = `Albumin [g/L]`,
    discontinued_medication = tolower(`Discontinued medication`) != "no",
    discontinued_medication_reason = if_else(discontinued_medication, `Discontinued medication`, NA_character_),
    best_supportive_care_from = if_else(tolower(`Best supportive care from:`) == "no", lubridate::as_date(NA), 
                                        janitor::excel_numeric_to_date(as.numeric(
                                          if_else(tolower(`Best supportive care from:`) == "no", NA_character_, `Best supportive care from:`) # Inner if_else to avoid warning
                                          ))),
    last_record = lubridate::as_date(`Date of last record`),
    outcome = factor(tolower(`Outcome (Discharged/Hospitalized/Transferred/Death)`), levels = tolower(outcome_labels), labels = outcome_labels),
    transferred_to = `Transferred to`
  )

@@

```{r}
progression_data_raw <- readxl::read_excel(file, sheet = "Disease Progression", range = "A1:CU2539", na = c("NA",""), col_types = c(c("text","text", "skip", "text", "text"), rep("text", 94)), col_names = TRUE)

progression_data <- progression_data_raw %>% 
  rename(patient_id = `Patient ID`, indicator = Indicator, first_day = First_Day, unit = Unit) %>%
  filter(!is.na(indicator))

@@

breathing_data <- progression_data %>% filter(indicator == "Breathing") %>%
  mutate(breathing = factor(tolower(value), levels = tolower(c("AA","Oxygen", "NIPPV","MV","ECMO")),labels = c("AA","Oxygen", "NIPPV","MV","ECMO"), ordered = TRUE))

breathing_unrecognized <- breathing_data %>% filter(is.na(breathing) != is.na(value))
if(nrow(breathing_unrecognized) > 0) {
  stop("Breathing NAs check failed")
}

@@

marker_map <- c("PCR" = "pcr_value", 
                "Supp. O2" = "oxygen_flow",
                "SpO2" = "SpO2",
                "SpO2 native" = "SpO2_native",
                "Horowitz index" = "Horowitz_index",
                "PEEP" = "PEEP",
                "Ferritin" = "ferritin",
                "D-dimer" = "d_dimer",
                "CRP" = "CRP",
                "IL-6" = "IL_6",
                "Lymphocyte count" = "lymphocyte_count",
                "Leukocyte count" = "leukocyte_count",
                "Leukocytes" = "leukocyte_count",
                "Leukocyte" = "leukocyte_count",
                "Hydroxychloroquine" = "hcq",
                "Azithromycin" = "az",
                "Kaletra" = "kaletra",
                "Tocilizumab" = "tocilizumab",
                "Amoxicillin/Clavulanate" = "amoxiclav",
                "Piperacilin/tazobactam" = "piperacilin_tazobactam",
                "Clarithromycin" = "clarithromycin",
                "ofloxacin" = "ofloxacin",
                "ceftriaxone" = "ceftriaxone",
                "cefuroxim" = "cefuroxim",
                "cefotaxim" = "cefotaxim",
                "cefepim" = "cefepime",
                "cefepime" = "cefepime",
                "Vancomycin" = "vancomycin",
                "ampicillin/sulbactam" = "ampicillin_sulbactam",
                "zinkorot" = "zinc",
                "sulfamethoxazol" = "sulfamethoxazol",
                "trimethoprim" = "trimethoprim",
                "Meropenem" = "meropenem",
                "Fluconazol" = "fluconazol",
                "Covalescent plasma" = "convalescent_plasma",
                "Anidulafungin" = "anidulafungin"
                )
#Check NAs (TODO)
unrecgonized_markers <- marker_data %>% filter(!(tolower(marker) %in% tolower(names(marker_map))))

@@

# Handle NT for SpO2_native
marker_data <- marker_data %>% 
  mutate(value = if_else(marker == "SpO2_native" & value == "NT", "< 92", value) )

# Some special cases
# TODO ask at hospitel
if(hospital_id == "QKuFp") {
  marker_data <- marker_data %>% mutate(value = if_else(marker == "SpO2" & value == "NEG", NA_character_, value))
}

>>>>

Notes to resolve:
- Missing "Covid medication before hospitalization" in "Old" version.

```{r}
i <- 5
file_relative <- raw_data_map$File[i]
hospital_id <- raw_data_map$HospitalID[i]
lang <- raw_data_map$Lang[i]
file_version <- raw_data_map$Version[i]

@@

if(file_version == "From Sheet") {
  file_version <- readxl::read_excel(file, sheet = version_sheet(lang), range = "C8:C8", col_types = "text", col_names = "version")$version 
}

if(!grepl("^(Old|1\\.[2-4]\\.[0-9])$", file_version)) {
  print(file_version)
  stop("Invalid version")
}

if(file_version == "Old") {
  n_middle_text_columns <- 12
  last_column <- "Z"
} else {
  n_middle_text_columns <- 13
  last_column <- "AA"
}

patient_col_types <- c("text","numeric","text","date","date", #ID - Date of admission
                                   rep("text", n_middle_text_columns), #Transferred from - Smoking
                                   rep("numeric", 5), # BMI - Albumin
                                   "text", "text", "date", "text") #Discontinued medication - Outcome

patients_range <- paste0("A2:",last_column, "100")

patient_data_raw <- readxl::read_excel(file, sheet = patient_sheet(lang), range = patients_range, na = "NA", col_types = patient_col_types) %>% 
  translate_patient_columns(lang)

patient_data_raw <- patient_data_raw %>%
  mutate(`Outcome (Discharged/Hospitalized/Transferred/Death)` = translate_outcomes(`Outcome (Discharged/Hospitalized/Transferred/Death)`, lang))

if(file_version == "Old") {
  patient_data_raw <- patient_data_raw %>% mutate(`Covid Medications before admission` = NA_character_)
}
```

```{r}

@@

logical_column <- function(x) {
  if(any( !is.na(x) && !(tolower(x) %in% c(yes_variants, no_variants))) ) {
    stop("Invalid logical column contents")
  }
  tolower(x) %in% yes_variants
}

@@

patient_data_typed <- patient_data_raw_usable %>% 
  transmute(
    hospital_id = hospital_id, 
    patient_id = `Patient ID`,
    age = Age, 
    sex = factor(Sex, levels = c("M","F")), 
    symptom_onset = lubridate::as_date(`Date of symptom onset`), 
    admission = lubridate::as_date(`Date of admission`), 
    transferred_from = `Transferred from`, 
    admitted_for_covid = logical_column(`Admitted for Covid`), 
    ischemic_heart_disease = logical_column(`Ischemic Heart Disease`),
    has_hypertension_drugs = !(tolower(`Hypertension drugs`) %in% no_variants),
    n_hypertension_drugs = if_else(has_hypertension_drugs,
                                   as.integer(if_else(has_hypertension_drugs, `Hypertension drugs`, NA_character_)), #Inner if_else to avoid warning
                                    0L),
    heart_failure = logical_column(`Heart failure`),
    COPD = logical_column(COPD),
    asthma = logical_column(`Asthma Bronchiale`),
    other_lung_disease = logical_column(`Other lung disease`),
    diabetes = logical_column(Diabetes),
    renal_disease = logical_column(`Renal Disease`),
    liver_disease = logical_column(`Liver Disease`),
    smoking = logical_column(Smoking),
    BMI = BMI,
    NYHA = NYHA,
    creatinin = `Creatinin [μmol/L]`,
    pt_inr = `PT INR (Quick)`,
    albumin = `Albumin [g/L]`,
    discontinued_medication = !(tolower(`Discontinued medication`) %in% no_variants),
    discontinued_medication_reason = if_else(discontinued_medication, `Discontinued medication`, NA_character_),
    best_supportive_care_from = if_else(tolower(`Best supportive care from:`) %in% no_variants, lubridate::as_date(NA), 
                                        janitor::excel_numeric_to_date(as.numeric(
                                          if_else(tolower(`Best supportive care from:`) %in% no_variants, NA_character_, `Best supportive care from:`) # Inner if_else to avoid warning
                                          ))),
    last_record = lubridate::as_date(`Date of last record`),
    outcome = factor(tolower(`Outcome (Discharged/Hospitalized/Transferred/Death)`), levels = tolower(outcome_labels), labels = outcome_labels)
  )

@@

```{r}
progression_data_raw <- readxl::read_excel(file, sheet = progression_sheet(lang), range = "A1:CU2539", na = c("NA",""), col_types = c(c("text","text", "skip", "text", "text"), rep("text", 94)), col_names = TRUE)

progression_data <- progression_data_raw %>% 
  translate_progression_columns(lang) %>%
  rename(patient_id = `Patient ID`, indicator = Indicator, first_day = First_Day, unit = Unit) %>%
  filter(!is.na(indicator)) %>%
  mutate(indicator = translate_markers(indicator, lang))

@@

breathing_data <- progression_data %>% filter(indicator == "Breathing") %>%
  mutate(
    # Special cases
    value = case_when(value == "Oxygen/NIPPV" ~ "NIPPV",
                      TRUE ~ value),
    # Translate to factor
    breathing = factor(tolower(value), levels = tolower(c("AA","Oxygen", "NIPPV","MV","ECMO")),labels = c("AA","Oxygen", "NIPPV","MV","ECMO"), ordered = TRUE))

breathing_unrecognized <- breathing_data %>% filter(is.na(breathing) != is.na(value))
if(nrow(breathing_unrecognized) > 0) {
  print(unique(breathing_unrecognized$value))
  stop("Breathing NAs check failed")
}

@@

marker_map <- c("PCR" = "pcr_value", 
                "Supp. O2" = "oxygen_flow",
                "SpO2" = "SpO2",
                "SpO2 native" = "SpO2_native",
                "Horowitz index" = "Horowitz_index",
                "PEEP" = "PEEP",
                "Ferritin" = "ferritin",
                "D-dimer" = "d_dimer",
                "CRP" = "CRP",
                "IL-6" = "IL_6",
                "PCT" = "procalcitonin",
                "Lymphocyte count" = "lymphocyte_count",
                "Leukocyte count" = "leukocyte_count",
                "Leukocytes" = "leukocyte_count",
                "Leukocyte" = "leukocyte_count",
                "Hydroxychloroquine" = "hcq",
                "Azithromycin" = "az",
                "Kaletra" = "kaletra",
                "Tocilizumab" = "tocilizumab",
                "Amoxicillin/Clavulanate" = "amoxiclav",
                "Piperacilin/tazobactam" = "piperacilin_tazobactam",
                "Clarithromycin" = "clarithromycin",
                "ofloxacin" = "ofloxacin",
                "ceftriaxone" = "ceftriaxone",
                "cefuroxim" = "cefuroxim",
                "cefotaxim" = "cefotaxim",
                "cefepim" = "cefepime",
                "cefepime" = "cefepime",
                "Vancomycin" = "vancomycin",
                "ampicillin/sulbactam" = "ampicillin_sulbactam",
                "zinkorot" = "zinc",
                "sulfamethoxazol" = "sulfamethoxazol",
                "trimethoprim" = "trimethoprim",
                "Meropenem" = "meropenem",
                "Fluconazol" = "fluconazol",
                "Covalescent plasma" = "convalescent_plasma",
                "Anidulafungin" = "anidulafungin",
                "Isoprinosine" = "isoprinosine"
                )
#Check NAs (TODO)
unrecgonized_markers <- marker_data %>% filter(!(tolower(marker) %in% tolower(names(marker_map))))

@@

# Handle NT for SpO2_native
marker_data <- marker_data %>% 
  mutate(value = if_else(marker == "SpO2_native" & tolower(value) == "nt", "< 92", value) )

# Some special cases
# TODO ask at hospital
if(hospital_id == "QKuFp") {
  marker_data <- marker_data %>% mutate(value = if_else(marker == "SpO2" & value == "NEG", NA_character_, value))
} else if(hospital_id == "YqNbe") {
  marker_data <- marker_data %>% mutate(value = if_else(marker %in% c("hcq","az") & value == "ano", ">100", value))
}
